Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-04-05 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis assumes that high-level musical concepts are agnostic and invariant to the unique tonal quality of the waveform, much like how a sheet music composition can be performed in a variety of ways using different instruments and styles while still retaining its musical essence.

We rely on a flexible and adaptable blueprint for musical expression, allowing performers to creatively interpret and express the same musical concepts in different ways.

Different waveforms can express the same musical concepts despite having distinct timbres or production styles, but it is crucial to distinguish between low-level and high-level musical features when analyzing or comparing different musical pieces. High-level features, such as melody, harmony, rhythm, tempo, form, and expression, are more abstract and can be recognized regardless of the specific waveform or production style used.

The thesis relies on the objective and comprehensive perspective on the musical content and meaning of a piece, much like reading sheet music can reveal the underlying musical structure and intention behind a piece of music.

One sheet music composition can be performed and recorded in a limitless variety of ways using different instruments, voices, tempos, and interpretations, resulting in an infinite number of valid waveforms that represent its musical content.

Conversely, the same waveform may have only a limited number of sheet music representations that accurately can capture its musical elements as per the current tradition and conventions. This is because the unique tonal quality of the waveform is strongly influenced by the specific instrument or production technology used to create it, making it difficult to transcribe its sonic properties into a precise sheet music notation. Moreover, this unique tonal quality is represented by high-level natural language concepts belonging to an old and rich cultural and sociological tradition.

>>>>

This thesis assumes that high-level musical concepts are agnostic and invariant to the unique tonal quality of the waveform, much like how a sheet music composition can be performed in infinite ways using different instruments and styles while retaining its musical essence.

We rely on a flexible and adaptable blueprint for musical expression, allowing performers to interpret and express the same musical concepts differently creatively.

Different waveforms can express the same musical concepts despite having distinct timbres or production styles. Still, it is crucial to distinguish between low-level and high-level musical features when analyzing or comparing different musical pieces. High-level features, such as melody, harmony, rhythm, tempo, form, and expression, are more abstract and can be recognized regardless of the specific waveform or production style.

The thesis relies on the objective and comprehensive perspective on a piece's musical content and meaning, much like reading sheet music can reveal the underlying musical structure and intention behind a piece of music.

One sheet music composition can be performed and recorded in a limitless variety of ways using different instruments, voices, tempos, and interpretations, resulting in an infinite number of valid waveforms representing its musical content.

Conversely, the same waveform may have only a limited number of sheet music representations that can accurately capture its musical elements as per the current tradition and conventions. This is because the unique tonal quality of the waveform is strongly influenced by the specific instrument or production technology used to create it, making it difficult to transcribe its sonic properties into a precise sheet music notation. Moreover, this unique tonal quality is represented by high-level natural language concepts belonging to a rich old cultural and sociological tradition.

### 2023-04-05 methods/subsections/model architecture and training strategy.tex
Updates from Overleaf

\section{Model architecture and training strategy}
Piaget's theory of cognitive development \cite{Piagets_Theory_of_Cognitive_Development} posits that children acquire knowledge from sensory and motor experiences in their early stages, from birth to around 18 months. This period referred to as the 'sensorimotor stage', sees the emergence of early representational thought through basic actions such as sucking, grasping, looking, and listening. As children progress through different developmental stages until adolescence and adulthood, their reasoning progressively moves toward abstract ideas and the use of deductive logic. Schemas, higher-order cognitive structures hypothesized to underlie many aspects of human knowledge and skill, emerge during this process, which Piaget interprets through an equilibration mechanism that balances new information with old knowledge. This mechanism involves \textit{assimilation} and \textit{accommodation}, which entail taking in new information to fit pre-existing schemas and modifying existing schemas resulting from new information. In the context of learning, complex structures are based on simpler ones, and transfer of learning occurs when dynamic structures capable of generalization are created. \cite{audioselfsupsurvey}

There are similarities between assimilation and accommodation in Piaget's theory of cognitive development and backpropagation in machine learning, both of which involve modifying existing structures based on new information. However, it should be noted that these processes' underlying mechanisms and contexts are distinct. While assimilation and accommodation involve modifying mental structures to incorporate new information, backpropagation entails adjusting neural network weights to improve output predictions. Assimilation and accommodation occur in cognitive development, while backpropagation is specific to machine learning. Thus, while there are broad similarities between these processes, they are fundamentally different and cannot be directly compared in detail.

>>>>

\section{Model architecture and training strategy}
Piaget's theory of cognitive development \cite{Piagets_Theory_of_Cognitive_Development} posits that children acquire knowledge from sensory and motor experiences in their early stages, from birth to around 18 months. This period, referred to as the \textit{sensorimotor stage}, sees the emergence of early representational thought through basic actions such as sucking, grasping, looking, and listening. As children progress through different developmental stages until adolescence and adulthood, their reasoning progressively moves toward abstract ideas and deductive logic. Schemas, higher-order cognitive structures hypothesized to underlie many aspects of human knowledge and skill, emerge during this process, which Piaget interprets through an equilibration mechanism that balances new information with old knowledge. This mechanism involves \textit{assimilation} and \textit{accommodation}, which entail taking in new information to fit pre-existing schemas and modifying existing schemas resulting from new information. In the context of learning, complex structures are based on simpler ones, and transfer of learning occurs when dynamic structures capable of generalization are created. \cite{audioselfsupsurvey}

There are similarities between assimilation and accommodation in Piaget's theory of cognitive development and backpropagation in DL, both of which involve modifying existing structures based on new information. However, it should be noted that these processes' underlying mechanisms and contexts are distinct. While assimilation and accommodation involve modifying mental structures to incorporate new information, backpropagation entails adjusting neural network weights to improve output predictions. Assimilation and accommodation occur in cognitive development, while backpropagation is specific to machine learning. Thus, while these processes have broad similarities, they are fundamentally different and cannot be directly compared in detail.
